%| Weekly report template for CSUS Senior Design
%|
%| language: LaTeX
%| Author: Ben Smith
%| 
%| This source has been tagged with the "<CHANGE" tag in areas
%| that require updating when making a new docuent
%|
%| This source will generate a PDF file complete with thumbnails navigation menu and metadata.
%| Much of the tex awesomeness comes from http://www.michaelshell.org/ praise be to him for creating the guide
\title{Lab Six: Special Projects}
\author{Ben Smith}

\input{./../../ProjectWideTex/header.tex}
\begin{document}

  \maketitle
    \begin{abstract}
      A number of projects are offered for a student to choose for a capstone project for a introductory logic course. These examples leverage a number of offerings from the Altera University Program tutorials for the DE0-Nano development board.
    \end{abstract}

  %| =================================================================================================
  %| Introduction
  %| =================================================================================================
  \section{\bfseries Introduction}
    \PARstart{T}{his} Lab allows the student to exercise some design liberty. A number of example projects will be provided but you are free to choose something of your own. The student could implement a communication protocol to one of the on board devices like the accelerometer.

    \subsection{Using paramaterized modules}
      Often a designer will not be required to reinvent the wheel during development. ``IP'' or intellectual property can be used to expediate design. These modules can be very complex like full processors or simple like a shift register, they can be very expensive or free of charge. Altera provides a number of paramterized modules free of charge with \href{ftp://ftp.altera.com/up/pub/Altera_Material/13.0/Tutorials/Verilog/Using_Library_Modules.pdf}{a tutorial on how to use them}.

  %| =================================================================================================
  %| Procedure
  %| =================================================================================================
  \section{\bfseries Lab Procedure}
    There is no specified procedure for this lab, It is acceptable to work alone or in a group.

    \subsection{Using the on board ADC}
    \subsection{Using the on board Accelerometer}
    \subsection{Implement a NIOSII softprocessor}

  %| =================================================================================================
  %| Lab Report Requirements
  %| =================================================================================================
  \section{\bfseries Lab Report}
    
  %| =================================================================================================
  %| Conclusions
  %| =================================================================================================
  \section{\bfseries Conclusion}
   
  %| =================================================================================================
  %| Bibliography
  %| =================================================================================================
  %\bibliographystyle{IEEEtran}
  %\bibliography{IEEEfull}
\end{document}